Finding finite transformations

### 2021-04-28 alpha/Schwarzian/schwarzian.tex
Finish note on Schwarzian and finite transforms

\vspace*{-.5\baselineskip}

@@

@@

	To find $S = S[w]$, consider \textbf{two successive transformations:}
	\begin{equation}
		z \xmapsto{\ w\ } w \xmapsto{\ v\ } v
	\end{equation}
	Here we've abused the symbols $w,v$ to label both the transformations and their images: $w\colon z\mapsto w(z)$, $v\colon w\mapsto v = v(w) = v(w(z))$. 
	To say that \textbf{$S[w]_{(z)}$ compose nicely} under successive transformations is to say that $S[w]_{(z)} = \Bqty{w,z}$ satisfies the following \textbf{chain rule:}
	\begin{equation}
		\Bqty{v,z} \dd{z}^2
		= \Bqty{v,w} \dd{w}^2
		+ \Bqty{w,z} \dd{z}^2
	\end{equation}
\pagebreak

@@

	Therefore, the infinitesimal transformation is saying that $X$ transforms in a non-trivial manner under Weyl rescaling:
	\begin{equation}
		\var_\omega g_{ab} = 2\var{\omega} g_{ab},
	\quad
		\var_\omega X
		= - \frac{\alpha' V}{2}\, 2\var{\omega}
	\end{equation}
	Actually, we have a similar result for the stress tensor transformation discussed above, which is associated with Weyl anomaly:
	\begin{equation}
		\var_\omega T(z)
		= - \frac{c}{12}\,\pd^2 (2\var{\omega})
	\end{equation}

	In fact, from $\var_\omega {X}$ we can already integrate over $\omega$ and recover the finite transformation for $\var_\omega {X}$, using the finite version of $
		2\var{\omega}
		= \pd\epsilon
			+ \pdbar\bar{\epsilon}
	$, namely the conformal Killing equation \eqref{eq:conformal_Killing}:
	\begin{equation}
		g'_{ab}
		= g_{ab}
		= e^{2\omega} \pdv{\sigma^c}{\sigma'^a}
			\pdv{\sigma^d}{\sigma'^b}\,g_{cd},
	\quad
		e^{-2\omega}
		= \frac{1}{d}\,
			g^{ab}
			\pdv{\sigma^c}{\sigma'^a}
			\pdv{\sigma^d}{\sigma'^b}
			\,g_{cd}
	\end{equation}
	For $d = 2$, note that $g_{z\bar{z}} = g_{\bar{z}z} = 2$, after carefully summing over indices, we have:
	\begin{equation}
		e^{-2\omega}
		= \frac{1}{2}\, g^{w\bar{w}}
			\pdv{z}{w}
			\pdv{\bar{z}}{\bar{w}}\,
			g_{z\bar{z}} \cdot 2
		= \pdv{z}{w} \pdv{\bar{z}}{\bar{w}},
	\quad
		e^{2\omega}
		= \pdv{w}{z} \pdv{\bar{w}}{\bar{z}}
		= \abs{\pd w}^2,
	\end{equation}
	\vspace{-.5\baselineskip}
	\begin{equation}
		X'(w)
		= X(z) - \frac{\alpha' V}{2}\, 2\omega
		= X(z) - \frac{\alpha' V}{2}\, \ln e^{2\omega}
		= X(z) - \frac{\alpha' V}{2}\, \ln \abs{\pd w}^2
	\end{equation}
\subsection{Worldsheet Dilaton Action}
	We see that under Weyl rescaling, $X$ gets a $X$-independent shift, proportional to the Weyl exponent. What does this mean physically? In fact, this is a ``remnant'' of the worldsheet dilaton term in bosonic string theory \cite{Polchinski:1998rq}:
	\begin{equation}
		S^\Phi = \frac{1}{4\pi\alpha'}
			\int \dd[2]{\sigma} \sqrt{g}\,
				\alpha' R\,\Phi(X)
	\end{equation}
	Although this term vanishes in the flat limit $R \to 0$, its effects persist in the stress tensor, since the stress tensor depends on the functional derivative of $g^{ab}$, and this does not vanish when $R \to 0$ \cite{Tong:2009np}:
	\begin{equation}
		T^\Phi_{ab}
		= -4\pi \frac{1}{\sqrt{g}} \fdv{S^\Phi}{g^{ab}}
		= \pqty{
				\cdv{a} \cdv{b} {}
				- g_{ab} \nabla^c \nabla_c
			}\,\Phi
		\xrightarrow{\ g_{ab} = \delta_{ab}\ }
		\pqty{
				\pd_a \pd_b
				- \delta_{ab}\, \pd^c \pd_c
			}\,\Phi
	\end{equation}

	Note that $T_{ab}$ is not traceless, namely the dilaton coupling $S^\Phi$ is generally \textbf{not Weyl invariant} at the classical level, unless $\pd^c \pd_c \Phi = 0$, or $\Phi(X) \equiv \Phi_0$ constant. 
	However, note that $R\,\Phi(X)$ comes with an $\alpha'$ in the Lagrangian; therefore, it's contribution mixes with $\order{\alpha'}$ loop corrections from the other terms, and it's possible that the \textbf{total combined Weyl anomaly cancels} in the presence of a non-trivial $\Phi(X)$, thus restoring Weyl invariance.

	In fact, the full theory is Weyl invariant if the \textit{dilaton profile} $\Phi(X)$, the spacetime metric $G_{\mu\nu}(X)$ and gauge field $B_{\mu\nu}(X)$ all have vanishing beta functions:
	\begin{equation}
		\beta^G_{\mu\nu}
		= \beta^B_{\mu\nu}
		= \beta^\Phi = 0
	\label{eq:beta_constraints}
	\end{equation}
	Which gives the spacetime equations of motion for $G_{\mu\nu}(X), B_{\mu\nu}(X)$ and $\Phi(X)$. 
	Suppose $B_{\mu\nu}(X) \equiv 0$; at 1-loop, we have:
	\begin{equation}
		0 = \beta^\Phi
		= \frac{D - 26}{6}
			- \frac{\alpha'}{2} \nabla^\mu \cdv{\mu} \Phi
			+ \alpha' \nabla^\mu \Phi \cdv{\mu} \Phi
			+ \order{\alpha'^2},
	\quad
		B_{\mu\nu}(X) \equiv 0
	\end{equation}
	We see that it's actually possible to move away from the critical dimension $D = 26$ by turning on a large gradient of $\Phi$, while keeping the theory consistent. In particular, the \textit{linear dilaton theory} is an \textit{exact} solution of \eqref{eq:beta_constraints} to \textit{all loops} given by:
	\begin{equation}
		G_{\mu\nu}(X) = \eta_{\mu\nu},
	\quad
		B_{\mu\nu}(X) = 0,
	\quad
		\Phi(X) = V_\mu X^\mu,
	\quad
		V^2 = \frac{26 - D}{6\alpha'}
	\end{equation}
	\vspace{-.5\baselineskip}
	\begin{equation}
		T(z)
		= T^X + T^\Phi
		= -\frac{1}{\alpha'}\, \normorder{\pd X \pd X}
			+ V \pd^2 X
	\end{equation}

	Finally, by computing $TX$ OPE and applying Ward identity, we can recover the transformation rule \eqref{eq:dilaton_transformation}.

%	TODO: repeat integration program
%	restrict to the holomorphic part of the transformation.

@@

\vspace{1.2\baselineskip}
%\pagebreak[4]
\raggedright
\printbibliography[%
%	title = {参考文献} %
	,heading = bibintoc
]
\end{document}

>>>>

@@

\setlength{\parskip}{.1\baselineskip}
\tableofcontents
\setlength{\parskip}{\parskipnorm}

@@

	To find $S = S[w]$, consider \textbf{two successive transformations:}
	\begin{equation}
		z \xmapsto{\ w\ } w \xmapsto{\ v\ } v
	\end{equation}
	Here we've abused the symbols $w,v$ to label both the transformations and their images: $w\colon z\mapsto w(z)$, $v\colon w\mapsto v = v(w) = v(w(z))$. 
	To say that \textbf{$S[w]_{(z)}$ compose nicely} under successive transformations is to say that $S[w]_{(z)} = \Bqty{w,z}$ satisfies the following \textbf{chain rule:}
	\begin{equation}
		\Bqty{v,z} \dd{z}^2
		= \Bqty{v,w} \dd{w}^2
		+ \Bqty{w,z} \dd{z}^2
	\label{eq:schwarzian_chain}
	\end{equation}
%\pagebreak

@@

	Therefore, the infinitesimal transformation is saying that $X$ transforms in a non-trivial manner under Weyl rescaling:
	\begin{equation}
		\var_\omega g_{ab} = 2\var{\omega} g_{ab},
	\quad
		\var_\omega X
		= - \frac{\alpha' V}{2}\, 2\var{\omega}
	\end{equation}
	Actually, we have a similar result for the stress tensor transformation discussed above, which is associated with \textbf{Weyl anomaly}:
	\begin{equation}
		\var_\omega T(z)
		= - \frac{c}{12}\,\pd^2 (2\var{\omega})
	\end{equation}

	In fact, from $\var_\omega {X}$ we can already integrate over $\omega$ and recover the finite transformation for $\var_\omega {X}$, using the finite version of $
		2\var{\omega}
		= \pd\epsilon
			+ \pdbar\bar{\epsilon}
	$, namely the conformal Killing equation \eqref{eq:conformal_Killing}:
	\begin{equation}
		g'_{ab}
		= g_{ab}
		= e^{2\omega} \pdv{\sigma^c}{\sigma'^a}
			\pdv{\sigma^d}{\sigma'^b}\,g_{cd},
	\quad
		e^{-2\omega}
		= \frac{1}{d}\,
			g^{ab}
			\pdv{\sigma^c}{\sigma'^a}
			\pdv{\sigma^d}{\sigma'^b}
			\,g_{cd}
	\end{equation}
	For $d = 2$, note that $g_{z\bar{z}} = g_{\bar{z}z} = 2$, after carefully summing over indices, we have:
	\begin{equation}
		e^{-2\omega}
		= \frac{1}{2}\, g^{w\bar{w}}
			\pdv{z}{w}
			\pdv{\bar{z}}{\bar{w}}\,
			g_{z\bar{z}} \cdot 2
		= \pdv{z}{w} \pdv{\bar{z}}{\bar{w}},
	\quad
		e^{2\omega}
		= \pdv{w}{z} \pdv{\bar{w}}{\bar{z}}
		= \abs{\pd w}^2,
	\end{equation}
	\vspace{-.5\baselineskip}
	\begin{equation}
		X'(w)
		= X(z) - \frac{\alpha' V}{2}\, 2\omega
		= X(z) - \frac{\alpha' V}{2}\, \ln e^{2\omega}
		= X(z) - \frac{\alpha' V}{2}\, \ln \abs{\pd w}^2
	\label{eq:finite_Weyl}
	\end{equation}
\subsection{Worldsheet Dilaton Action}
	We see that under Weyl rescaling, $X$ gets an $X$-independent shift, proportional to the Weyl exponent. What does this mean physically? In fact, this is a ``remnant'' of the worldsheet dilaton term in bosonic string theory \cite{Polchinski:1998rq}:
	\begin{equation}
		S^\Phi = \frac{1}{4\pi\alpha'}
			\int \dd[2]{\sigma} \sqrt{g}\,
				\alpha' R\,\Phi(X)
	\end{equation}
	Although this term vanishes in the flat limit $R \to 0$, its effects persist in the stress tensor, since the stress tensor depends on the functional derivative of $g^{ab}$, and this does not vanish when $R \to 0$ \cite{Tong:2009np}:
	\begin{equation}
		T^\Phi_{ab}
		= -4\pi \frac{1}{\sqrt{g}} \fdv{S^\Phi}{g^{ab}}
		= \pqty{
				\cdv{a} \cdv{b} {}
				- g_{ab} \nabla^c \nabla_c
			}\,\Phi
		\xrightarrow{\ g_{ab} = \delta_{ab}\ }
		\pqty{
				\pd_a \pd_b
				- \delta_{ab}\, \pd^c \pd_c
			}\,\Phi
	\end{equation}

	Note that $T_{ab}$ is not traceless, namely the dilaton coupling $S^\Phi$ is generally \textbf{not Weyl invariant} at the classical level, unless $\pd^c \pd_c \Phi = 0$, or $\Phi(X) \equiv \Phi_0$ constant. 
	However, note that $R\,\Phi(X)$ comes with an $\alpha'$ in the Lagrangian; therefore, its contribution mixes with $\order{\alpha'}$ loop corrections from the other terms, and it's possible that the \textbf{total combined Weyl anomaly cancels} in the presence of a non-trivial $\Phi(X)$, thus restoring Weyl invariance.

	In fact, the full theory is Weyl invariant if the \textit{dilaton profile} $\Phi(X)$, the spacetime metric $G_{\mu\nu}(X)$ and gauge field $B_{\mu\nu}(X)$ all have vanishing beta functions:
	\begin{equation}
		\beta^G_{\mu\nu}
		= \beta^B_{\mu\nu}
		= \beta^\Phi = 0
	\label{eq:beta_constraints}
	\end{equation}
	Which gives the spacetime equations of motion for $G_{\mu\nu}(X), B_{\mu\nu}(X)$ and $\Phi(X)$. 
	Suppose $B_{\mu\nu}(X) \equiv 0$; at 1-loop, we have:
	\begin{equation}
		0 = \beta^\Phi
		= \frac{D - 26}{6}
			- \frac{\alpha'}{2} \nabla^\mu \cdv{\mu} \Phi
			+ \alpha' \nabla^\mu \Phi \cdv{\mu} \Phi
			+ \order{\alpha'^2},
	\quad
		B_{\mu\nu}(X) \equiv 0
	\end{equation}
	We see that it's actually possible to move away from the critical dimension $D = 26$ by turning on a large gradient of $\Phi$, while keeping the theory consistent.

	In particular, the \textit{linear dilaton theory} is an \textit{exact} solution of \eqref{eq:beta_constraints} to \textit{all loops} given by:
	\begin{equation}
		G_{\mu\nu}(X) = \eta_{\mu\nu},
	\quad
		B_{\mu\nu}(X) = 0,
	\quad
		\Phi(X) = V_\mu X^\mu,
	\quad
		V^2 = \frac{26 - D}{6\alpha'}
	\end{equation}
	\vspace{-.5\baselineskip}
	\begin{equation}
		T(z)
		= T^X_{zz} + T^\Phi_{zz}
		= -\frac{1}{\alpha'}\,\normorder{\,\pd X \pd X}
			+ V \pd^2 X
	\end{equation}
	The reason that this holds to \textit{all loops} is that we can \textit{define} an exact CFT with this action and stress tensor $T(z)$. Although $T^\Phi_{ab}$ is generally not traceless, in this case its trace vanishes on-shell: $\pd \pdbar X = 0$, so indeed we have conformal invariance.

	Finally, by computing $TX$ OPE and applying Ward identity, we \textbf{basically \textit{impose} the transformation rule \eqref{eq:dilaton_transformation} on the field $X$}. One can thus interpret $\var_\omega X$ as some sort of \textbf{anomalous quantum ``dimension'' of $X$}; in this case it is not a rescaling of $X$ like what we usually encounter, but is instead a \textit{shift}, introduced to preserve Weyl (conformal) invariance in the quantum theory.

\subsection{Finite Transformation of Linear Dilaton}
	Although we've already obtained the finite transformation \eqref{eq:finite_Weyl} by viewing it as a Weyl rescaling, we would like to repeat the same program for finding finite transformations, just like what we've done for the Schwarzian derivative, and then confirm the above results.

	In the case of linear dilaton, there are no indices of worldsheet coordinates in $X$, so the composition of finite transformations $F[w]_{(z)}$ simply gives the following ``chain rule'':
	\begin{equation}
		F[v]_{(z)} = F[v]_{(w)} + F[w]_{(z)}
	\end{equation}
	There is no extra $\dd{z}^2$ factors like the ones in the Schwarzian chain rule \eqref{eq:schwarzian_chain}.

	For simplicity, we shall restrict to the \textbf{holomorphic part} of the transformation.
	We note that $F$ converts the product in the usual chain rule $
		\dv{v}{z}
		= \dv{v}{w} \dv{w}{z} 
	$ into a sum. This in fact implies that $F$ must be proportional to $\log \pd w$. But let's be patient and again consider the flow $w_t(z)$ as in the Schwarzian case; we now have:
	\begin{equation}
		\pdd{t} F
		= - \frac{\alpha'V}{2}
				\pdv{w_t} \pdd{t} w_t
		= - \frac{\alpha'V}{2}
				\frac{1}{\pd w_t} \pdd{t} \pd w_t
		= - \frac{\alpha'V}{2}
				\pdd{t} \ln \pd w_t
	\end{equation}
	Indeed we have $
		F = - \frac{\alpha'V}{2} \ln \pd w
	$, as expected. If we include the anti-holomorphic part, we recover the finite transformation \eqref{eq:finite_Weyl}:
	\begin{equation}
		X'(w)
		= X(z) - \frac{\alpha' V}{2}\, \ln \abs{\pd w}^2
	\end{equation}

@@

%\vspace{1.2\baselineskip}
%\pagebreak[4]
\raggedright
\printbibliography[%
%	title = {参考文献} %
	,heading = bibintoc
]
\end{document}
